---
title: "lab9x"
author: "Marisa Abundis"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```


#### **Conditional statements with if-else**

* if structure
```r
if(<condition>) {
        ## do something
} 
## Continue with rest of code
```

* if/else structure
```r
if(<condition>) {
        ## do something
} 
else {
        ## do something else
}
```

* if and else if structure - series of tests
```r
if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}
```
* Valid if/else structure
```{r}
## Generate a uniform random number
x <- runif(1, 0, 10)  
if(x > 3) {
        y <- 10
} else {
        y <- 0
}
```


#### **for Loops**

* used to iterate over elements of an object (list, vector, etc)
```{r}
for(i in 1:10) {
        print(i)
}
```

```{r}
x <- c("a", "b", "c", "d")
for(i in 1:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

* seq_along() function + for loops are used to genereate an interger seequence based on the length of an object
```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```

* Not necessary to use index-type variable
```{r}
for(letter in x) {
        print(letter)
}
```


* for one line loops, curly brackers not neccessary
```{r}
for(i in 1:4) print(x[i])
```


#### **Nested for loops**

* used for multidimensional or hierarchial data structures (matrices & lists)
```{r}
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```


#### **While loops**

* runs loop while conidition is true
```{r}
count <- 0
while(count < 10) {
        print(count)
        count <- count + 1
}
```

* Multiple conditional statements
```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```


#### **Functions in R**

* simple function that does nothing
```{r}
f <- function() {
        ## This is an empty function
}
## Functions have their own class
class(f)  
```

```{r}
## Execute this function
f() 
```

* function that determines how many times the string is printed
```{r}
f <- function() {
        cat("Hello, world!\n")
}
f()
```

* function that retursn total num of characters printed to the console
```{r}
f <- function(num) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(3)
```

```{r}
print(meaningoflife)
```
* set default value for num
```{r}
f <- function(num = 1) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
f()    ## Use default value for 'num'
```

```{r}
f(2)   ## Use user-specified value
```

```{r}
f(num = 2)
```

Functions: 

* has one formal argument named num with a default value of 1. The formal arguments are the arguments included in the function definition. The formals() function returns a list of all the formal arguments of a function
